---
title: "Group Project"
date: "4/15/2021"

output:
  pdf_document:
    number_sections: no
  html_notebook:
    number_sections: no
  html_document:
    number_sections: no
  word_document: default

---

```{r setup, include=FALSE}
library("haven")    
library("AER")   
library("fastDummies")
library("plm")
df <- read.csv("table1.csv")
head(df)

#generate dummies
df$dummy8090 <- ifelse(df$Year < 1990, 1, 0)
df$dummy9000 <- ifelse(df$Year >= 1990 & df$Year < 2000, 1, 0)
df$dummy0010 <- ifelse(df$Year >= 2000 & df$Year < 2010, 1, 0)
df$dummy1018 <- ifelse(df$Year >= 2010 , 1, 0)

df <- dummy_cols(df, select_columns = "State")

# try to involve state dummies 
cols <- 9:59
nval <- 0:50
df$stateid <- t(sweep(df[, cols], 2, nval, "*"))[t(df[, cols]) != 0]
df$stateid

```

# Introduction

This project aims to analyze how changes in state minimum wages affect changes in the state unemployment level, and we have built three model to explore this relationship. In the first basic model, we simply regress state’s unemployment rate on effective minimum wage in 2020 dollars. In the second model, based on the first basic model, we consider year fixed effect and state fixed effect, and we also introduce three dummy variables for 1980-1990, 1990-2000, and 2000-2010 to check whether or not the results change across the three different periods. In the third model, we build a panel regression model in order to include state fixed effect and time fixed effect. Our analysis indicates that state minimum wages have a positive relationship with state unemployment rate.

# Result
## Model 1: baseline model
```{r, echo = T, warning= F}
# model1:
fm1 <- lm(unemployment_rate ~ Effective.Minimum.Wage.2020.Dollars , data = df)
summary(fm1)
coeftest(fm1, df = Inf, vcov = vcovHC(fm1, type = "HC1"))

```
### Interpretation 1
If minimum wage increases by 1 dollar, then unemployment rate would increase by 0.635777% / 63.58 percentage points. (not sure which guys please check it !!!!!!!)


## Model2: 
base model+year fixed effect 

### Goal
in order to account of time trend in unemployment rate, we included time dummies in our model
```{r, echo = T, warning= F}
#model2: base model+year fixed effect 
fm2 <- lm(unemployment_rate ~ Effective.Minimum.Wage.2020.Dollars+ 
            dummy9000 + dummy0010 + dummy1018, data = df)
summary(fm2)
coeftest(fm2, df = Inf, vcov = vcovHC(fm2, type = "HC1"))

```
### interpretation 2
After accounting for time trend, we had a similar finding: if minimum wage increases by 1 dollar, then unemployment rate would increase by 0.474070% / 47.41 percentage points. (not sure which guys please check it !!!!!!!)


## Model 3:
base model+year fixed effect + state fixed effect 
### Goal
In this model, we included state dummies because geographic factors might be associated with unemployment rate.

```{r, echo = T, warning= F}
#model3: base model+year fixed effect + state fixed effect 
fm3 <- lm(unemployment_rate ~ Effective.Minimum.Wage.2020.Dollars+ 
            factor(stateid) + 
            dummy9000 + dummy0010 + dummy1018, data = df)
summary(fm3)
coeftest(fm3, df = Inf, vcov = vcovHC(fm3, type = "HC1"))

```
### interpretation 3
After accounting for time and state dummies, we had a similar finding: if minimum wage increases by 1 dollar, then unemployment rate would increase by 0.43830% / 43.83 percentage points. (not sure which guys please check it !!!!!!!)


## Model 4
base model+year fixed effect + year $\cdot$ effective + state fixed effect + state $\cdot$effective

### Goal
In this model, we aim to explore the association between minimum wage and unemployment rate across year periods

```{r, echo = T, warning= F}
#model4: base model+year fixed effect + year*effective + state fixed effect + state*effective
fm4 <- lm(unemployment_rate ~ Effective.Minimum.Wage.2020.Dollars+ 
            Effective.Minimum.Wage.2020.Dollars*dummy9000+
            Effective.Minimum.Wage.2020.Dollars*dummy0010+
            Effective.Minimum.Wage.2020.Dollars*dummy1018+
            factor(stateid) + 
            dummy9000 + dummy0010 + dummy1018, data = df)
summary(fm4)
coeftest(fm4, df = Inf, vcov = vcovHC(fm4, type = "HC1"))

```
### Interpretation 4
we found that minimum wage was asscoiated with increased unemployment rate, and we found that compared to baseline time periord (1980-1990) if minimum wage increases by 1 dollar, then unemployment rate would decrease by -102.5792 percentage points (1990-2000), -43.9424 percentage points (2000-2010), -87. 5224 percentage points for (2010-2018). (not sure which % or percentage points  guys please check it !!!!!!!)





